---
title: "Assessment 1"
author: "Angelina Navitskaya, s4647285"
date: "10/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RMarkdown Editing

[***RMarkdown***](https://rmarkdown.rstudio.com/lesson-8.html) represents *a unified authoring framework for data science, combining code, its results, and user's prose commentary*. ***RMarkdown*** documents are written in **an easy-to-write plain text format and contain chunks of embedded R code**. Documents are fully reproducible and support dozens of output formats including PDFs, Word files, slideshows, and more.

-----------------------------------------------------------
-----------------------------------------------------------

Q.6 Insert an R chunk and create a variable with this dataset
https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv. The name of the variable should be `fastfood`.

```{r}
library(tidyverse)
fastfood <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv")
```

Q.7 Display the first 10 rows of the dataset using head() and kable(). 

```{r echo= FALSE, results='asis'}
library(knitr)
head(fastfood, 10)

kable(fastfood[1:10, ], caption = "First 10 rows")
```

Display the first 10 rows of the dataset and the first 5 variables

```{r echo= FALSE, results='asis'}
library(knitr)
kable(fastfood[1:10,1:5])
```

Q.8 Display the observations that has more than 1000 calories.

```{r}
fastfood %>%
  filter(calories > 1000)
```

Q.9 Arrange observations with more than 40 in total_fat and more than 80 in total_carb in the descending order and save them to a new variable (dataset) called `dont_eat_this`

#Only 1 variable can be arranged in the descending order at a time. The assumption is that "in the descending order" goes after "total_carb" and therefore, this variable should be considered.

```{r}
library(dplyr)
dont_eat_this <- fastfood %>%
  filter(total_fat > 40,
         total_carb > 80) %>%
  arrange(desc(total_carb))
```

Q.10 Using the initial dataset variable, use `case_when()` to create a new variable `heavy_food` which is equal to "heavy" when total calories are greater than 500, "low" when total calories are less than 250 and "average" for all other cases. Count the number of observations in each created category.

```{r}
fastfood %>%
  mutate(heavy_food = case_when(calories > 500 ~ "heavy",
                                calories < 250 ~ "low",
                                T ~ "average")) %>%
  group_by(heavy_food) %>%
  count(heavy_food)
```

Q.11 Display the types of variables in the dataset using `skimr` package 

```{r}
#This dataset has 3 character and 15 numeric variables.

library(skimr)
skim(fastfood)
```

Q.12 Present the count observations from each restaurant in a descending order. Show the number of distinct items on a menu in the dataset.

```{r}
fastfood %>%
  count(restaurant) %>%
  arrange(desc(restaurant))

#There are 505 distinct items on a menu in the dataset.

n_distinct(fastfood$item)
```

Q.13 Using groupings (group_by()), summarise and display the average number of calories for each restaurant.

```{r}
#The average of all numbers in a set is MEAN.

fastfood %>%
  group_by(restaurant) %>%
  summarise(mean(calories))
```

Q.14 Add variables to the dataset, which:

-  calculates the average calories per type of restaurant and call it `average_calories` 
-  calculates the maximum total_fat per type of restaurant and call it `max_fat` 
-  calculates the minimum cholesterol per type of restaurant and call it `min_cholesterol` 

```{r}
fastfood %>%
  group_by(restaurant) %>%
  mutate(average_calories = mean(calories)) %>%
  mutate(max_fat = max(total_fat)) %>%
  mutate(min_cholesterol = min(cholesterol)) %>%
  select(restaurant, item, calories, total_fat, cholesterol, average_calories, max_fat, min_cholesterol)
```

Q. 15. Display the data vis of total fat per each type of restaurant. Write a narration (2-3 sentences) why you believe this type of data viz presents such information best. 

```{r}
#Total fat is a continuous variable, whereas restaurant is discrete. According to the ggplot cheat sheet, there are 4 data visualization types for this case scenario including geom_col, geom_boxplot, geom_dotplot and geom_violin. For this task geom_col was considered as the most appropriate as the heights of the bars best represent values in the data per each type of restaurant. It is simple and easy to understand.

fastfood %>%
  group_by(restaurant) %>%
  ggplot(aes(restaurant, total_fat)) +
  geom_col() +
  coord_flip() +
  labs(x = "Restaurant", y = "Total Fat", 
       title = "Total Fat Distribution Across Different Restaurants")
```

Q.16 Add a variable to the dataset, which calculates the sum of cholesterol and sodium and call it `cholesterol_sodium`.

```{r}
fastfood %>%
  mutate(cholesterol_sodium = cholesterol + sodium)
```

Remove the variable `salad`

```{r}
fastfood <- fastfood %>%
  select(-salad)
```

Q.17 Use observations for Mcdonalds to plot sugar variable against protein with `geom_point()` 

```{r}
fastfood %>%
  filter(restaurant == "Mcdonalds") %>%
  ggplot(aes(sugar, protein, colour = restaurant)) +
  geom_point() +
  labs(x = "Sugar",
       y = "Protein",
       colour = "Restaurant",
       title = "Comparison of Sugar Variable against Protein")
```

Q.18 Identify variable(s) which should be factors and transform their type into a factor variable.

```{r}
#Factor variables are categorical variables that can be either numeric or string variables having a limited number of different values. Therefore, the only variable that can be transformed into a factor is restaurant which has only 8 different values.

class(fastfood$restaurant)
#[1] "character"
fastfood$restaurant <- as_factor(fastfood$restaurant)
class(fastfood$restaurant)
#[1] "factor"
```

Q.19 Create two new variables:

Read about `cut_number()` function using Help and add a new variable to the dataset `calories_type`. Use `calories` variable for `cut_number()` function to split it into 3 categories `n=3`, add labels `labels=c("low", "med", "high")` and make the dataset ordered by arranging it according to calories. 

```{r}
fastfood <- fastfood %>%
  mutate(calories_type = cut_number(calories, n = 3, labels = c("low", "med", "high"))) %>%
  arrange(calories_type)
```

Q.20 Create a dataviz that shows the distribution of `calories_type` in food items for each type of restaurant. Think carefully about the choice of data viz. Use facets, coordinates and theme layers to make your data viz visually appealing and meaningful. Use factors related data viz functions.

```{r}
#Both variables "calories_type" and "restaurant" are discrete. 

fastfood %>%
  ggplot(aes(calories_type, restaurant, colour = restaurant)) +
  geom_count() +
  facet_wrap(~restaurant) +
  labs(title = "Distribution of Calories in Food Items for Each Restaurant")
```

Q.21 The second variable should show the percentage of `trans_fat` in `total_fat`. Add the variable to the dataset and call it `trans_fat_percent`. Do not forget to save the updated dataset.

```{r}
fastfood <- fastfood %>%
  mutate(trans_fat_percent = trans_fat / total_fat *100)
```

Q.22 Create a dataviz that shows the distribution of `trans_fat` in food items for each type of restaurant. Think carefully about the choice of data viz. Use facets, coordinates and theme layers to make your data viz visually appealing and meaningful.

```{r}
fastfood %>%
  ggplot(aes(restaurant, trans_fat)) +
  geom_count() +
  facet_wrap(~restaurant) +
  coord_flip() +
  labs(x = "Restaurant", y = "Trans Fat", 
       title = "Total Fat Distribution Across Different Restaurants") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 1))
 

```

Q.23 Calculate and show the average (mean)  `total_fat` for each type of restaurant. No need to save it as a variable.

```{r}
fastfood %>%
  group_by(restaurant)%>%
  summarise(mean(total_fat))
```

Q.24 And create a dataviz that allow to compare different restaurants on this variable (`total_fat`). You can present it on one dataviz (= no facets). 

```{r}
fastfood %>%
  group_by(restaurant) %>%
  ggplot(aes(restaurant, total_fat)) +
  geom_col() +
  coord_flip() +
  labs(x = "Restaurant", y = "Total Fat", 
       title = "Total Fat Distribution Across Different Restaurants")
```

